---
title: "PracticeR"
author: "Marion Geary"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Practice R

Welcome to R! As a little background, R is an open source statistical programming language written largely in C. It is largely a functional programming language. (There is some object oriented stuff but it's super weird.) This is an R Markdown file, which is pretty much R's version of a Jupyter Notebook combining markdown text and code.

To start, let's import some libraries.

```{r}
# This is a super popular collection of packages designed for data science. It provides standardization 
# across all of its packages which is super helpful (and not always the case with an opensource 
# language). When using the tidyverse, dataframes are called "tibbles".
library(tidyverse)
# This is a collections of packages for machine learning! It is created with the same principles and syntax
# as the tidyverse
library(tidymodels)
# This collection of packages has all of our data and functions. It mimics the syntax of the tidyverse
library(nflverse)
```

Press the green play button in the upper right hand corner to run the whole code chunk or run a single line using `Cmd/Ctrl + Enter`.

To add a code chunk, use `Cmd/Ctrl + Opt + I`.

The assignment operator in `R` is `<-`.

```{r}
# Load in data from the nflverse
player_stats <- load_player_stats(seasons = most_recent_season())
player_stats_kicker <- load_player_stats(seasons = most_recent_season(), stat_type = "kicking")
roster <- load_rosters(seasons = most_recent_season())
team_data <- load_teams()
```

Now we can work with the data! Check out these [help files](https://www.rstudio.com/resources/cheatsheets/) to get started - I recommend starting with `dplyr` for filtering data or `ggplot` for creating graphs.

Here's a quick run through of how to manipulate data.

```{r}
# The `%>%` operator is called a "pipe". It's a way to combine multiple functions in one line of code. Example:
player_stats_grouped <- group_by(player_stats, recent_team)
player_stats_grouped_max <- summarise(player_stats_grouped, max_fantasy_points = max(fantasy_points_ppr))

# Now let's do the same thing with a pipe %>%
max_fantasy_points_per_team <- player_stats %>% group_by(recent_team) %>% summarise(max_fantasy_points = max(fantasy_points_ppr))

# filter() gets rows
player_stats_filter <- player_stats %>% filter(week == 5)

# select() gets columns
player_stats_select <- player_stats %>% select(starts_with("rushing") | starts_with("receiving") | starts_with("player"))

# group_by() lets you manipulate data based on a group. The tibble looks the same, but it applies other functions
# to each group rather than to the whole at once
max_fantasy_points_per_team <- player_stats %>% group_by(recent_team) %>% summarise(max_fantasy_points = max(fantasy_points_ppr))

# summarise() creates a summary table
player_stats_summary <- player_stats %>% summarise(avg = mean(fantasy_points_ppr))

# mutate() changes the tibble. For example, create a new column
player_stats_mutate <- player_stats %>% mutate(percent_completion = completions / attempts)
```

We can also make lots of graphs! Check out the `ggplot` cheat sheet for more details.

The basic structure is `ggplot(data = <DATA>) + <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))`

```{r}
# Here are some plot examples
ggplot(player_stats) + geom_point(aes(x = attempts, y = completions, color = season_type))

ggplot(player_stats) + geom_boxplot(aes(x = fantasy_points_ppr)) + facet_wrap(~ recent_team)

ggplot(player_stats %>% filter(player_name == "M.Stafford" | player_name == "J.Burrow"), mapping = aes(x = week, y = fantasy_points_ppr)) + geom_point() + facet_wrap(~ player_name) + geom_smooth() + labs(x = "Week", y = "Fantasy Points")
```

We can also integrate Python with R! Create a python code chunk:

```{python}
for i in range(0, 5):
  print(i)
```

Notice that this calls the package `reticulate` from the `tidyverse`. The cheatsheet is very helpful. Here is a closer look at some helpful Python-in-R functions.

```{python}
# We can access r objects in python! And manipulate them!
# Use r.obj_name
r.player_stats.head()
panthers_df = r.player_stats[r.player_stats["recent_team"] == 'CAR']
panthers_df.tail()
```

```{r}
# this should already be installed in the tidyverse
# Use py$obj_name
library(reticulate)
# We can access python objects in r too
panthers_r <- py$panthers_df %>% filter(week == 5)

# Run a single line of python
# The code needs to be passed as a string
py_run_string("print(r.panthers_r['fantasy_points'].mean())")
```

I recommend checking out the [nflverse documentation](https://nflverse.nflverse.com/) - all of the packages in this library have some cool functions that are worth exploring, and the documenation is pretty comprehensive.

Have fun exploring! Feel free to change this file.
